---
title: "R Notebook"
output: html_notebook
---

## Load in packages 
First let's load in the packages that we'll use in our analysis.  These packages provide us with tools and additional functionality that will make it easier to conduct our analysis.
```{r}
library(tidyverse)
library(tidycensus)
```


```{r}
census_api_key("a26529040c6be62df861c3beb8159a03068ba77e")
```



<br/>
https://api.census.gov/data/2010/dec/sf1/variables.html
```{r}
vars10 <- c("P005003", "P005004", "P005006", "P004003")
NY_race <- get_decennial(geography = "county", variables = vars10, year = 2010,
                    summary_var = "P001001", state = "NY")
```



```{r}
NY_race <- NY_race %>% 
  spread(key = "variable", value = "value")
```

```{r}
NY_race <- NY_race %>% 
  rename(County = "NAME",
         County_population = "summary_value",
         County_White = "P005003",
         County_Black_AfAm = "P005004",
         County_Asian = "P005006",
         County_Hispanic_Latino = "P004003") %>% 
  
  separate(County, into = c("County", "State"),
           sep = ",") %>% 
  
  mutate(County = str_replace(County, "County",""))
```


```{r}
NY_race <- NY_race %>% 
  mutate(County = str_trim(County, side = "both"))
```



```{r}
NY_income_zip <- get_acs(geography = "zcta", variables = "B19013_001",
                     year = 2012)
```

```{r}
NY_income_zip <- NY_income_zip %>% 
  separate(NAME, 
           into = c("code","School_Zip_Code"),
           sep = " ") %>% 
  select(School_Zip_Code, Median_HH_income = estimate)
```




## Load in the data and get familiar with the contents
Today we are going to use data on lead (Pb) in drinking water in New York State schools.  The data is from the New York State Department of Health and can be [accessed here](https://health.data.ny.gov/Health/Lead-Testing-in-School-Drinking-Water-Sampling-and/rkyy-fsv9).  

Now let's load in the data that we will use in our analysis.  

The code below loads in a **csv** file (essentially an Excel file) with the data and converts it into a data table in R.  
```{r}
Pb_data <- read_csv("https://github.com/stahlm/Hackathon_2020/raw/master/data/Lead_Testing_in_School_Drinking_Water_Sampling_hackathon.csv")
```

```{r}
Pb_data <- Pb_data %>% 
  mutate(County = str_trim(County, side = "both")) %>% 
  mutate(County = if_else(County == "Saint Lawrence", "St. Lawrence", 
                          County)) %>% 
  mutate(School_Zip_Code = as.character(School_Zip_Code))
```


```{r}
Pb_data <- left_join(Pb_data, NY_race,
                     by = "County")
```

```{r}
Pb_data <- left_join(Pb_data, NY_income_zip,
                     by = "School_Zip_Code")
```

```{r}
Pb_data <- Pb_data %>% 
  select(School,
         Type_of_Organization,
         County, 
         School_District, 
         School_City,
         School_Zip_Code,
         School_Street,
         Number_of_Outlets_LT_15_ppb,
         Number_of_Outlets_GT_15_ppb,
         Any_Buildings_with_Lead_Free_Plumbing,
         Out_of_Service,
         County_Location,
         Median_HH_income,
         County_population,
         County_White,
         County_Black_AfAm,
         County_Asian,
         County_Hispanic_Latino
         )
```


```{r}
Pb_data <- Pb_data %>% 
  mutate(County_Location = str_replace(County_Location,"\\(" ,"")) %>% 
  mutate(County_Location = str_replace(County_Location,"\\)" ,"")) %>% 
  separate(County_Location, 
           into = c("Latitude", "Longitude"),
           sep = ",")
```


```{r}
write_csv(Pb_data, 
          "./data/Lead_schools_hackathon_dataset.csv")
```

